Verification of a TCP/IP stack

### 2021-05-18 SecDev21/paper.tex
shortening the information sections

TCP is a connection-oriented protocol; a
connection between the sender and the receiver must be established before
transmitting any data. Furthermore, TCP is a reliable protocol since it
guarantees the delivery of all the messages and that they are delivered in the
order they were sent. It also provides error-checking mechanisms to discard and
recover any corrupted data.

@@

@@

Generally, every TCP communication session will go through the following three phases (if no error occurs): 1)~opening the connection 2)~sending and receiving
the data and 3)~closing the connection. The behavior of the three-phase communication can be described by the state machine in \cref{Fig:statemachine}.

An edge represents a state transition. A transition is either triggered by an
explicit action, stated in italics format on the label's edge, for example,
\textit{Close}, or is triggered by the arrival of a specific flag/s.
An edge-label in the format of $\frac{\ \ x \ \ }{\ \ y \ \ }$
represents the associated flags transmitted with each transition, where $x$
is either send or receive and $y$ refers to the actual
flag/s transmitted in response to the action that caused the state-transition.
An explicit action is either a user-triggered action (through the user's API)
or an automatically performed action triggered by a timeout event. The flags are
embedded in the header section of a transmitted segment and can be one of the
following:
\begin{itemize}
\item \ack: the last message received by the sender is acknowledged.
\item \syn: this flag is sent to establish a connection.
\item \fin: this flag is sent to close the connection.
\item \rst: this flag is sent to reset the connection, often when an error
occurred in one or the other side.
\end{itemize}

The state machine in figure~\ref{Fig:statemachine} does not represent the
complete protocol specification; for example, it does not reflect
error-conditions or any actions which are not connected with the state changes.
Rather, it gives an overview of all the possible states a TCP connection could
reach over its lifetime.

Each state has a meaning. \slisten represents waiting for a connection request
from any remote TCP and port. \ssynsent and \ssynrcv represent the
initialization of the connection, that ends when \sestab is reached.
\sfwone, \sfwtwo, \sclosing and \stimewait are a group of states that represents
waiting for a connection termination request from remote TCP. \sclosew and
\slastack represent waiting for a connection termination request from local
user. Finally, \sclosed represents no connection state at all.

\subsection{TCP Multitasking model}

Different tasks can interact and update the socket data structure to handle the
various events that can occur within a TCP session. The TCP norm, under
Section “3.9. Event Processing” page 52, describes a possible implementation
of how to handle these events based on three tasks: one for the
\emph{user calls}, one for the \emph{arriving segments} and one for the
\emph{timers}.
The role of each of the three tasks can be summarized as follow:
User calls refer to functions, namely \texttt{open}, \texttt{close},
\texttt{abort}, \texttt{send} and \texttt{receive}, that can be called
by the user to control the connection, send, or receive data.
These functions can trigger transitions between the connection's states since
they are intended to control the connection.
In the arriving segments task, the received segments are being processed,
and the corresponding messages are sent back. Transitions between
states can be triggered on the reception of a segment.
Finally, timers control the timeouts, for example, the retransmission
timeout to retransmit a message, or the time-wait timeout to close the
connection after a specified amount of time elapsed. Thus, corresponding
transitions to the timeout events can also be triggered by this task.

\section{CycloneTCP library}
\label{sec:stack}

CycloneTCP is a professional-grade embedded TCP/IP library developed by the
Oryx Embedded company. It provides several network protocols whose implementations
are meant to conform with the Request for Comments (RFC) Internet standards.
Figure~\ref{Fig:TcpStack} shows the protocols implemented in the library.
These include the IPv4 or Ipv6 of the network layer, TCP and UDP of the transport
layer and, among other, DHCP and HTTP of the application layer.
In particular, the TCP implementation is conform with the RFC 793
norm~\cite{rfc793}.

The library is written in ANSI C, and it supports a large number of 32-bits
embedded processors and a large number of Real-time Operating System (RTOS).
Hence, a model based on multiple tasks has been adopted to implement the TCP
protocol as described in the norm, with, in addition, complex cooperation
mechanisms. A task is dedicated to the user code, another to the timers and the
last is assigned to the processing of the incoming segments.
It can also run on bare metal environment.

In the CycloneTCP implementation, the socket data structure is used to retain
the status of a TCP connection. Furthermore, a socket can be
manipulated by the user through the TCP's library Application Programming
Interface (API) to perform various TCP operations, such as the transmission of
data.

The quality of the C implementation hailed in the AMNESIA:33 report~\cite{AMNESIA33}
as one of the most resilient TCP/IP stack. This gives a really good assurance
in the code and it strengthens the confidence that we can have in a formally
verified code running inside the library environment.
\todo{Add anything else relevant about the CycloneTCP library}.

>>>>

@@

TCP is a connection-oriented protocol; a connection between the sender and the receiver must be established before transmitting any data. Furthermore, TCP is a reliable protocol since it guarantees the delivery of all the messages and that they are delivered in the order, they were sent. It also provides error-checking mechanisms to discard and recover any corrupted data.

@@

Generally, every TCP communication session will go through the following three phases (if no error occurs): 1)~opening the connection, 2)~sending and receiving the data, and 3)~closing the connection. The behaviour of the three-phase communication can be described by the state machine given in \cref{Fig:statemachine}. An edge represents a state transition. A transition is either triggered by an explicit action, stated in italics format on the label's edge, for example, \textit{Close}, or is triggered by the arrival of a specific flag/s. An edge-label in the format of $\frac{\ \ x \ \ }{\ \ y \ \ }$ represents the associated flags transmitted with each transition, where $x$ is either send or receive and $y$ refers to the actual flag/s transmitted in response to the action that caused the state transition. An explicit action is either a user-triggered action (through the user's API) or an automatically performed action triggered by a timeout event. The flags are embedded in the header section of a transmitted segment and can be one of the following:
\begin{itemize}
\item \ack: the last message received by the sender is acknowledged.
\item \syn: this flag is sent to establish a connection.
\item \fin: this flag is sent to close the connection.
\item \rst: this flag is sent to reset the connection, often when an error
occurred on one or the other side.
\end{itemize}

The state machine in \cref{Fig:statemachine} does not represent the complete protocol specification; for example, it does not reflect error conditions or any actions which are not connected with the state changes. Instead, it gives an overview of all the possible states a TCP connection could reach over its lifetime. Each state has a meaning. \slisten represents waiting for a connection request from any remote TCP and port. \ssynsent and \ssynrcv represent the initialization of the connection that ends when \sestab is reached. \sfwone, \sfwtwo, \sclosing, and \stimewait are a group of states that represents waiting for a connection termination request from remote TCP. \sclosew and \slastack represent waiting for a connection termination request from the local user. Finally, \sclosed represents no connection state at all.

%\subsection{TCP Multitasking model}

TCP is based on a multitasking model. Different tasks can interact and update the socket data structure, used to retain the status of a TCP connection, to handle the various events that can occur within a TCP session. The TCP norm, \cite{rfc793} , under Section "3.9. Event Processing" page 52, describes a possible implementation of how to handle these events based on three tasks: one for the \emph{user calls}, one for the \emph{arriving segments} and one for the \emph{timers}. The role of each of the three tasks can be summarized as follow: User calls refer to functions, namely \texttt{open}, \texttt{close}, \texttt{abort}, \texttt{send} and \texttt{receive}, that can be called by the user to control the connection, send, or receive data. These functions can trigger transitions between the connection's states since they are intended to control the connection. In the arriving segments task, the received segments are being processed,
and the corresponding messages are sent back. Transitions between states can be triggered on the reception of a segment. Finally, timers control the timeouts, for example, the retransmission timeout to retransmit a message or the time-wait timeout to close the connection after a specified amount of time elapsed. Thus, corresponding transitions to the timeout events can also be triggered by this task. A complete description of the protocol's specification can be found at the \emph{RFC 793} document \cite{rfc793}.

The CycloneTCP implementation adopts the three-task-based model and the socket data structure. Thus, one task is dedicated to the user's code, one to the timers, and one to the processing of the incoming segments. The user can manipulate a socket through the TCP's library Application Programming Interface (API) to perform various TCP operations, such as data transmission.
